---
title: "HHID Exploration.Rmd"
date: "6/22/2021"
output: html_document
---


```{r setup, include=FALSE}
library(rmarkdown)
library(tidyverse)
library(retroharmonize)


knitr::opts_chunk$set(echo = TRUE)


```

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 300px;
}
```
# An exploration of Household IDs across years in PHL LFS data 

First, load the metadata files
```{r, echo=FALSE}
# load metadata
load("Y:/GLD-Harmonization/551206_TM/PHL/PHL_data/I2D2/Rdata/metadata.Rdata")

```

## Seeing if HHIDs are truly panel or unique across rounds

First I want to know if our initial assumption -- that we should append all four quarters -- is valid. Are HHIDs consistent within a year, between rounds/quarters? Let's see.

### Which years have the `intvw` variable?

Some years, but not all, have a factor variable that tells us if the household was in the previous quarter. Let's use the metadata file to take a look.

```{r}
# identify "intvw" variable across years 
intv <- metadata %>%
  filter( grepl("intvw", var_name_orig)) %>%
  filter(!is.na(labels)) %>%
  select(id, labels, label_orig) %>%
  unnest_longer(labels, 
                values_to = "value",
                indices_to = "value_label") 

intv
```

It looks like years 01-04, 06, 07, 12, and 13 have this variable -- but not all rounds in those years. It looks like the only years that actually count for replacement or repeat vists are years 01-03. However, another *potential* renamed value for this variable is `panel`, which appears in years 97-00 with values 1-4. This doesn't present a perfect match though, since if you go into the actual `.dta` files you find out that you have values for each of the 1-4 values in each of the rounds, which doesn't make sense if the variable is a "placeholder" for the round/panel indicator.

```{r}
metadata %>%
  filter( grepl("panel", var_name_orig)) %>%
  select(id, labels, label_orig, class_orig) %>%
  unnest_longer(labels, 
                values_to = "value",
                indices_to = "value_label") 

```



## What are the actual counts of unique HH as I have them coded now?

Taken from the `.do` file info

| year | count of Unique HH | observations |
| ------| ----------------| ------------------|
| 1997 | 39,274| 202,742|
| 1998 | 39,837| 207,663|
| 1999 | 38,830| 200,902|
| 2000 | 39,273| 202,738|
| 2001 | 39,716| 207,457|
| 2002 | 41,947| 793,795|
| 2003 | .| .|
| 2004 | .| .|
| 2005 | .| .|
| 2006 | .| .|
| 2007 | .| .|
| 2008 | 159,161| 797,746|
| 2009 | 159,740| 804,166|
| 2010 | 163,744| 795,851|
| 2011 | 165,473| 807,631|
| 2012 | 92,031| 817,049|
| 2013 | 93,981| 818,118|
| 2014 | 84,485| 805,307|
| 2015 | 107,658| 820,811|
| 2016 | 107,658| 820,811|
| 2017 | .| .|
| 2018 | 10,632| 722,237|
| 2019 | 160,309| 707,022|

