---
title: "renv: A Quick Guide"
output: 
  github_document:
    html_preview: false
    toc: true
    toc_depth: 4
editor_options: 
  markdown: 
    wrap: 72
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = here::here("Support/Guides and Documentation"),
      envir = globalenv()
    )
  })
---

A Quick guide for using `renv` . Please refer to the [getting started
vignette](https://rstudio.github.io/renv/articles/renv.html) for more.

# What is `renv`?

A detailed explanation of `renv` can be [found
here](https://rstudio.github.io/renv/articles/renv.html) but, in short,
it is a way to standardize and keep track of R packages across users.

This standardization and transparency is important in the context of
reproducibility: internally, all GLD harmonizers are using the same
versions of R packages and new users can run the code with exactly the
same versions of the R packages that produced the results. This is only
one method we use to ensure our results are reproducible; see [this
explanation for
more](https://rstudio.github.io/renv/articles/renv.html#reproducibility-1).

# How We Use `renv`

While there are many ways to collaborate with `renv` our workflow is set
up with the following tools assumed:

1.  Version control using the [GLD GitHub
    Repository](https://github.com/worldbank/gld). You can, of course,
    clone it and make your own repository and it will work the same.

2.  `.Rproject` file. This file stores a few lines of repository-level
    preferences. `renv` is associated with this `.Rproject` file, and is
    available on the GitHub repository.

3.  `renv.lock` file. This file contains a list of all the package
    names, versions, and information used in the current version of the
    repository — but not the package files themselves. This file is also
    version-controlled on the repository.

## Using `renv` for the first time

Since this `renv` project is already initialized, first-time users will
need to only

1.  Open Rstudio with the `gld.Rproj` file

2.  Install the `renv` package

    ```{r, eval=FALSE}
    install.packages("renv")
    ```

3.  Finally, tell `renv` about the `renv.lock` file, which will install
    all packages. This may take a while for the first time.

    ```{r, eval=FALSE}
    renv::restore()
    ```

If loaded correctly, the console will tell you that the project `gld` is
loaded with renv.

## Using `renv` daily

1.  Open Rstudio with the `gld` project file — it will automatically
    load `renv`.

### When I update or install a package

The beauty of `renv` is if one user finds an update or new package for
the project, it will alert the others *somewhat* automatically. But you
have to tell it to do so. First install/update your new package

```{r, eval=FALSE}
install.packages("a brand new package!")
```

Then tell `renv` to save the state of your library. You can also do this
in the UI under "Packages" then `renv`

```{r, eval=FALSE}
renv::snapshot()
```

This updates `renv.lock` file. If you're keeping track of this file on
GitHub as we are, commit this change to tell others.

### Getting Updates

You see that there's a new commit on your version control; the
`renv.lock` has been updated in your branch. Pull the update. Then,
either from the UI or in the console,

```{r, eval=FALSE}
renv::restore()
```

This will install the packages that are missing or have been updated.

# Help! (light)

#### If you're not sure if `renv` is initialized on the branch you're using

For most cases, users will be coming to a branch with `renv` already
initialized. If you want aren't sure if you have `renv` initialized,
type

```{r, eval=FALSE}
renv::status()
```

#### If you need to initialize `renv` on your branch

If for some reason `renv` is not initialized on your branch, follow
steps 1 and 2 above in "Using renv for the First Time". Then initialize
using

```{r, eval=FALSE}
renv::init()
```

and select option 2 when it asks you about libraries.

#### If `renv` doesn't load

Then try

```{r, eval=FALSE}
renv::load()
```

#### `renv::snapshot()` doesn't detect new packages/updates

Try ensuring that it's capturing all types of snapshots.

```{r, eval=FALSE}
renv::settings$snapshot.type("all")
renv::snapshot()
```

#### 

```{r}
citation("renv")
```
