---
title: "PHL: Weights, Population, Labor Force Participation Checks and Methodology"
output: 
  html_document:
    html_preview: false
    toc: true
    toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = here::here("Support/Country Survey Details/PHL/LFS"),
      envir = globalenv()
    )
  })
---

```{r setup, include=FALSE}
# thanks to https://bookdown.org/yihui/rmarkdown-cookbook/hook-scroll.html, this will allow for output scrolling when
# max.height is set.
options(width = 80)
options(dplyr.summarise.inform = FALSE)

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
local({
  hook_output <-knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})

# declare top-level project directory
here::i_am("main.R")

```

```{r, include=FALSE}
# source main script
source(here::here("main.R"), local = knitr::knit_global())
```

## Introduction

This document outlines the methods for calculating the final weights
used the in GLD survey and reproduces steps -- along with outputs -- to
check that GLD figures align with those published by the Philippine
Statistics Authority (PSA).

Survey weights, expanded population, and labor force participation are
all fundamental concepts to the GLD that are inextricably linked:
quality handling of the weight variable will help ensure that the final
survey will generate accurate measurements of the country's population
and labor force figures of that year. This document is indented to
provide users with a step-by-step walkthrough of how we analyzed,
calculated, and checked these three figures. Since we use a script-based
approach instead of a notebook-based approach, this code is an adapted
version of [population_lfp.R](/Support/Global/population_lfp.R).

### A note on our data

The Philippines Labor Force Survey generally consists of four rounds per
year. The GLD data use all four rounds when available; however, for
years 1998-2001, we only have the first round of data (January).
Therefore, you will see below that for these years the appropriate
mathematical adjustments have been made to account for the fact that we
are only using one round. However, our final yearly indicators of
magnitude (i.e., population) in 1998-2001 will naturally be much lower
than the figures the PSA publishes since we only use about one quarter
of the data.

### External Sources

All PSA figures have been referred to at either:

-   the main Labor Force Survey page:
    <https://psa.gov.ph/statistics/survey/labor-and-employment/labor-force-survey>

-   or the ILO Comparison Annex file, (direct link):
    <https://psa.gov.ph/sites/default/files/attachments/hsd/pressrelease/Table%2057%20Philipine%20Concept%20vs.%20ILO%20Concept.pdf>

## Data Work

### Weights

#### Import the GLD Data

We load tidyverse and the previously appended GLD output data for all
survey years, 1997 - 2019.

```{r, echo=FALSE}
library(tidyverse)
load(file = file.path(PHL, "PHL_data/GLD/population.Rdata"))
```

#### Determine which years and rounds need adjustment

It turns out that not all of the weight variables are uniform in how
they scale the data by year or wave, even though the correct weight
variable was selected. This could throw off our results. So some scaling
will need to be done. Let's take a look.

```{r, max.height = '200px'}
sum.w.a <- phl %>%
  group_by(year, wave) %>%
  summarise(weight_median = median(weight, na.rm = TRUE),
            weight_mean   = mean(weight, na.rm = TRUE))

sum.w.a %>% print(n=Inf)
```

#### Adjust the weights accordingly

From working with the Philippines LFS data, we know that scaling down
these "high" weights by 10,000 typically puts the higher-valued weights
in line with the other weight values. Let's scale the appropriate years
and waves, then check.

Also, we'll create `partic` , a binary indicator for labor force
participation. We'll create it from the GLD variable `lstatus`: a
categorical variable that takes three non-missing values. In our case,
any non-missing value but `3` / `"Non-Labor Force"` will be coded as
`partic == TRUE` . Then, `3` / `"Non-Labor Force"` will be coded as
`partic == FALSE` . In the case of the Philippines, `lstatus` is coded
directly from variable data provided by the PSA; when available in the
survey, we always use the ILO definition of employment status.

```{r, echo=FALSE}
# create a weight-adjusted objected
phl2 <- phl %>%
  mutate(
    partic = case_when( (lstatus == 1 | lstatus == 2) ~ TRUE,
                        (lstatus == 3) ~ FALSE ),
    weight2= case_when( year <= 2004 ~ weight / 10000,
                        year == 2005 & (wave == "Q1" | wave == "Q2") ~ weight / 10000,
                        year == 2006 ~ weight / 10000,
                        year == 2007 & (wave == "Q1") ~ weight / 10000,
                        TRUE         ~ weight))  
```

#### Update the weights summary object

Now that we have `weight2` in our new `phl2` object that's adjusted,
tentatively, let's add a by-quarter summary of this new weight variable
and add it to the `sum.w.a`.

```{r, max.height = '200px'}
# create `weight_corrected` vector
weight.adj <- phl2 %>%
  group_by(year, wave) %>%
  summarise(weight_median_corrected = round(median(weight2, na.rm = TRUE)),
            weight_mean_corrected   = round(mean(weight2, na.rm = TRUE)))

weight.adj %>% print(n=Inf)
```

```{r, max.height = '200px'}
# join to original weight summary object
sum.w <- sum.w.a %>%
  left_join(weight.adj, by = c("year", "wave")) %>%
  mutate(weight_median_corrected = weight_median_corrected,
         weight_mean_corrected   = weight_mean_corrected)

sum.w %>% print(n=Inf)
```

Looks good. The scaling by `10,000` seemed to bring the weight variables
in line -- and we appeared to code the correct years and waves. Let's
move on to check our Population and Labor Force Participation Figures
with `weight2`

### Checks: Population and LFP

#### Total vs. age 15 and above

We could use the total population, but the PSA provides their figures
for ages 15 and above. So we'll make two versions, just in case.

```{r, max.height = '200px'}
# Whole population
sum.y <- phl2 %>%
  group_by(year) %>%
  summarize(
    pop = sum(weight2),
    lfp = weighted.mean(partic, weight2, na.rm = TRUE) * 100)

sum.y %>% print(n=Inf)
```

Note that we filter out all observations below the age of 15 for the
second object.

```{r, max.height = '200px'}
# for those 15 and over (inclusive)
sum.15.y <- phl2 %>%
  filter(age >= 15) %>%
  group_by(year) %>%
  summarize(
    pop_15up = sum(weight2),
    lfp_15up = weighted.mean(partic, weight2, na.rm = TRUE) * 100)

sum.15.y %>% print(n=Inf)
```

#### Adding PSA figures

The PSA data, linked above, we'll add simply by input in R. It's not
ideal, but it is reproducible should we find an error or point of
contention later on. This will allow us to compare the GLD figures for
population and labor force participation with those from the PSA.

On the PSA figures themselves, the PSA appear to publish yearly
summaries consistently after 2007, but only quarterly before. When the
PSA do publish a yearly figure, that figure is used. However, when only
quarterly figures are available, we aproximate the yearly figures via a
simple mean, recognizing that this is a just an estimation exercise. For
those following along with the PSA data, you'll notice the figures are
in reverse chronological order: years are descending and likewise
within-year figures read in reverse order. Visually the pattern follows

```{r, eval=FALSE}
# year == "2016" ~ [year],
# year == "2015" ~ mean(c([Q4], [Q3], [Q2], [Q1])),
```

We'll add six new variables. The first two, `psa_lfp_15up` and
`psa_pop_15up` give the PSA's figures for Labor Force Participation and
Population for those over 15. The second set, `dif_pop_15up` and
`dif_lfp_15up`, are difference variables that take the simple difference
between GLD and PSA figures. Finally, `flag_pop_5pct` and
`flag_lfp_2pct` are TRUE/FALSE or boolean variables that warn us if a
certain year's population error is more than 5% or the labor force
participation error is more than 2%.

```{r}
sum.15.y <- sum.15.y %>%
  mutate(
    psa_lfp_15up = case_when(
      year == "2019" ~ 61.3,
      year == "2018" ~ 60.9,
      year == "2017" ~ 61.2,
      year == "2016" ~ 63.4,
      year == "2015" ~ mean(c(63.3, 62.9, 64.6, 63.8)),
      year == "2014" ~ 64.4,
      year == "2013" ~ 63.9,
      year == "2012" ~ 64.2,
      year == "2011" ~ 64.6,
      year == "2010" ~ 64.1,
      year == "2009" ~ 64.0,
      year == "2008" ~ 63.6,
      year == "2007" ~ mean(c(63.2, 63.6, 64.5, 64.8)),
      year == "2006" ~ mean(c(63.8, 64.6, 64.8, 63.6)),
      year == "2005" ~ mean(c(64.8, 64.6, 64.8, 63.2)),
      year == "2004" ~ mean(c(63.8, 64.1, 65.4, 64.1)),
      year == "2003" ~ mean(c(64.5, 63.5, 64.0, 63.3)),
      year == "2002" ~ mean(c(63.9, 64.4, 66.3, 64.3)),
      year == "2001" ~ mean(c(62.6)), #only use Q1
      year == "2000" ~ mean(c(62.7)), #only use Q1
      year == "1999" ~ mean(c(63.1)), #only use Q1
      year == "1998" ~ mean(c(62.7)), #only use Q1
      year == "1997" ~ mean(c(63.1, 63.2, 65.5, 63.3))
    ),
    psa_pop_15up = case_when(
      year == "2019" ~ 72931000,
      year == "2018" ~ 71339000,
      year == "2017" ~ 69896000,
      year == "2016" ~ 68125000,
      year == "2015" ~ 66622000,
      year == "2014" ~ 62189000,
      year == "2013" ~ 61176000,
      year == "2012" ~ 62985000,
      year == "2011" ~ 61882000,
      year == "2010" ~ 60717000,
      year == "2009" ~ 59237000,
      year == "2008" ~ 57848000,
      year == "2007" ~ 56864000,
      year == "2006" ~ 55638000,
      year == "2005" ~ 54797000,
      year == "2004" ~ 53562000,
      year == "2003" ~ 52305000,
      year == "2002" ~ 50841000,
      year == "2001" ~ 48413000, #Q1 population data
      year == "2000" ~ 47185000, #Q1 population data
      year == "1999" ~ 45852000, #Q1 population data
      year == "1998" ~ 44517000, #Q1 population data
      year == "1997" ~ 44143000
    ),
    dif_pop_15up = pop_15up - psa_pop_15up,
    dif_lfp_15up = lfp_15up - psa_lfp_15up,
    err_pop_15up = (dif_pop_15up/pop_15up),
    err_lfp_15up = (dif_lfp_15up/lfp_15up),
    flag_pop_5pct= (err_pop_15up >= 0.05),
    flag_lfp_2pct= (err_lfp_15up >= 0.05)
  ) 

```

#### Final Notes

We see that our scaling method has produced a weight variable, `weight2`
that generates population and labor force figures that are very
consistent with those published by the PSA. All but two, or 21 of 23
years, pass our self-imposed consistency tests. Even so, the data for
these two years that do not pass are still within reasonable margins of
error. What is more, labor force participation rate for these same years
appears to be nearly identical with the PSA figures. The case of years
2013-2014 and the summary data as a whole suggest that any issues with
the weight variable are likely reflected in final measurements of
magnitude and more than ratios.

### Graphs

We can visualize our final results below. The graphical results
reiterate the finding that the differences between the population
figures of GLD and those from the PSA do not reflect discrepancies in
measurements of labor force participation.

```{r}
gg.pop <- ggplot(sum.15.y, aes(year, err_pop_15up)) +
  geom_col(aes(fill = flag_pop_5pct)) +
  scale_y_continuous(labels = scales::label_percent(), limits = c(0,0.2)) +
  labs(title = "Relative Error in GLD vs PSA Popupation 15yrs and Older",
       y = "Relative Error",
       x = "Year",
       fill= "Error >= 5%") +
  theme_minimal()

gg.pop
```

```{r}
gg.lfp <- ggplot(sum.15.y, aes(year, err_lfp_15up)) +
  geom_col(aes(fill = flag_lfp_2pct)) +
  scale_y_continuous(labels = scales::label_percent(), limits = c(-0.1, 0.2)) +
  labs(title = "Relative Error in GLD vs PSA Labor Force Participation 15yrs and Older",
       y = "Relative Error",
       x = "Year",
       fill= "Error >= 5%") +
  theme_minimal()

gg.lfp
```

#### 
